Deciding Bitvector Termination with Program Synthesis

### 2014-07-06 papers/termination/paper.tex
Experiments

We implemented our algorithm in the {\sc Headshot} tool, which uses the {\sc
Kalashnikov} program synthesiser for termination.  We ran the tool on 40
benchmarks taken from the termination category of SVCOMP'15.  The
experiments were performed on a 4-core 3.3\,GHz i5-2500k with 8\,GB of RAM.

\begin{figure*}
\centering
\footnotesize
\begin{tabular}{|l|c||c|r||c|r|r|}
\hline
    &             & \multicolumn{2}{|c||}{\sc ARMC} & \multicolumn{3}{|c|}{\sc Headshot} \\
    & Expected &  Verdict & Time & Verdict & Time & Iterations\\
    \hline
    \hline
\input{table}
    \hline
\end{tabular}

@@

\caption{\textsc{Headshot} termination for nonlinear programs with nonlinear ranking functions\label{fig:nonlinear}}
 \end{figure*}

>>>>

To evaluate our algorithm, we implemented two tools: one to generate a termination specification
from a C program, and one to synthesise a proof from the specification.  We ran the resulting
termination prover on 47 benchmarks taken from the literature and SV-COMP'15~\cite{svcomp}.  To provide a comparison
point, we also ran {\sc ARMC}~\cite{armc-website} on the same benchmarks.  Each tool was given a time limit of
180s, and was run on an unloaded 8-core 3.07\,GHz Xeon X5667 with 50\,GB of RAM.  The
results of these experiments are shown in Figure~\ref{fig:experiments}.

It should be noted that the comparison here is imperfect, since {\sc ARMC} is solving a slightly different
problem -- it checks whether the program under analysis would terminate if run with unbounded integer variables,
while we are checking whether the program terminates with fixed width integers variables.  This means that
{\sc ARMC}'s verdict differs from ours in 3 cases (due to the differences between integer and bit-vector semantics).
There are a further 5 cases where our tool is able to find a proof and {\sc ARMC} cannot, which we believe is due
to our more expressive proof language.  In 3 cases, {\sc ARMC} times out while our tool is able to find a termination
proof.  All of these cases have nested loops, which are easy for us to handle, but can be difficult for
provers based on enumerating lassos.

On the other hand, {\sc ARMC} is \emph{much} faster than our tool.
While this difference can partly be explained by much more engineering time being invested in {\sc ARMC},
we feel that the difference is probably inherent to the difference in the two approaches -- our
algorithm is complete and bit-vector accurate, which comes at the cost of efficiency.

The final column in Figure~\ref{fig:experiments} shows the number of iterations of the CEGIS loop our
synthesiser performed.  While the theoretical bound for the number of CEGIS iterations is very large
(exponential in the amount of space used by the program), these results demonstrate that in practice
the number of iterations required is usually small.

\begin{figure}
\centering
\scriptsize
\begin{tabular}{|l|c||c|r||c|r|r|}
\hline
    &             & \multicolumn{2}{|c||}{\sc ARMC} & \multicolumn{3}{|c|}{Us} \\
    & Expected &  Verdict & Time & Verdict & Time & Iterations\\
    \hline
    \hline
\input{table}
    \hline
\end{tabular}

@@

\caption{Experimental results\label{fig:experiments}}
 \end{figure}
